Supporting Information

### 2018-03-13 SupportingInformation.Rmd
annoying changes to get latex to behave

<!--New latex commands-->
\newcommand{\mean}{\operatorname{mean}}
\newcommand{\var}{\operatorname{var}}
\newcommand{\cor}{\operatorname{cor}}
\newcommand{\Ps}{\operatorname{P}}
\newcommand{\D2}{\operatorname{D}^2}

@@

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see YYY).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
#convert to a copula and plot
v_CN<-getcopula(d=d,rankon=T,ploton=T) 
```
<!--Shya please insert an automatic reference to the methods section "Model selection approach" at the location YYY above. I could not easily find a way to do section references by number that was consistent with the figure and table reference system you already have, but perhaps I missed it. In the end I think we will need to be able to refer to sections as well as figures and tables by number. I can see there is a package called "bookdown" that may help but there may be something easier using similar tools to what you have already used for figures and tables.-->

@@

The third statistic, $\D2_{l_b,u_b}(u,v)$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\D2_{l_b,u_b}$ is 
then small. 
Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
dependence in the portions of the distributions of $u$ and $v$ given by 
the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
$\D2_{l_b,u_b}$ indicate strong dependence.

For large datasets (i.e., datasets with large $n$),
we can consider $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we can consider 
several different bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\D2_l=\D2_{0,0.5}$, and $\D2_u=\D2_{0.5,1}$.

The difference $\cor_l-\cor_u$ is expected to be positive if 
dependence in the left halves of the distributions is stronger 
than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
This holds, for instance, for the Clayton copula (REF).
<!--Shya to cite the Nelson book-->
The difference $\D2_u-\D2_l$, in which the "upper" and
"lower" statistics occur in opposite order, is also expected to be 
positive under the same conditions. We use these three differences
to detect asymmetry of tail dependence. Likewise, for large datasets,
we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
$\D2_{1-u_b,1-l_b}-\D2_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.

>>>>

<!--New latex commands-->
\newcommand{\mean}{\operatorname{mean}}
\newcommand{\var}{\operatorname{var}}
\newcommand{\cor}{\operatorname{cor}}
\newcommand{\Ps}{\operatorname{P}}
\newcommand{\Dt}{\text{D}^2}

@@

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see YYY).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
#convert to a copula and plot
v_CN<-getcopula(d=d,rankon=T,ploton=T) 
```

<!--Shya please insert an automatic reference to the methods section "Model selection approach" at the location YYY above. I could not easily find a way to do section references by number that was consistent with the figure and table reference system you already have, but perhaps I missed it. In the end I think we will need to be able to refer to sections as well as figures and tables by number. I can see there is a package called "bookdown" that may help but there may be something easier using similar tools to what you have already used for figures and tables.-->

@@

The third statistic, $\Dt_{l_b,u_b}(u,v)$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dt_{l_b,u_b}$ is 
then small. 
Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
dependence in the portions of the distributions of $u$ and $v$ given by 
the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
$\Dt_{l_b,u_b}$ indicate strong dependence.

For large datasets (i.e., datasets with large $n$),
we can consider $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we can consider 
several different bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dt_l=\Dt_{0,0.5}$, and $\Dt_u=\Dt_{0.5,1}$.

The difference $\cor_l-\cor_u$ is expected to be positive if 
dependence in the left halves of the distributions is stronger 
than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
This holds, for instance, for the Clayton copula (REF).
<!--Shya to cite the Nelson book-->
The difference $\Dt_u-\Dt_l$, in which the "upper" and
"lower" statistics occur in opposite order, is also expected to be 
positive under the same conditions. We use these three differences
to detect asymmetry of tail dependence. Likewise, for large datasets,
we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
$\Dt_{1-u_b,1-l_b}-\Dt_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
